### 2024-12-03 paper/paper.qmd
Finished the introduction part of the paper, with the necessary citations

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Download the relevant packages needed for downloading required data if
# necessary. If packages are already downloaded, comment out the following lines.
install.packages("palmerpenguins")
install.packages("here")
install.packages("modelsummary")
install.packages("rstanarm")

# Load the necessary packages
library(tidyverse)
library(palmerpenguins)
library(here)
library(modelsummary)
library(rstanarm)
```

# Introduction

Canada provides shelters for people in need, such as refugees, the homeless, and people experiencing transitions in living spaces, and the country is continuing to build new shelters for such residents. On April 16, 2024, the Canadian government announced a funding of over 250 million dollars over two years to address homelessness and encampment all over the country. In 2022, Canada announced a target to eliminate homelessness by 2030. Canada Mortgage and Housing Corporation also allocated a total of 420 million dollars to account for homelessness (@ShelterInvestment). This implicates that the Canadian government will continuously invest in improving current shelters and building new shelters around the country. In 2024, Canada's Minister of Immigration, Refugees, and Citizenship Marc Miller also announced that the country would accept 27,000 new refguees this year, and this trend is likely to continue in the next years (@CanadaRefugee). We can conclude that more shelters are likely to be built in Canada in the future, and the same trend is also valid for the GTA(Greater Toronto Area).

To find a way to maximize efficiency of the investment on shelters, I want to evaluate the effect of predicting then number of beds (which reflects the necessary size of shelters need be built, expanded, or renovated) by constructing a multiple linear regression model based on variables such as specific types of shelters and their services, as well as occupancy rates and averaged number of occupants per day. We will limit our scope of prediction to the GTA(Greater Toronto Area), since it is the area around the largest city in Canada, and is a popular destination in Canada for refugees and homeless. Specifically, the aim of this paper is to just to just test the validity/effect of using such a statistical model to predict the number of beds, so we will set our sample for modelling to only cover Down Toronto. Downtown Toronto is relatively representative of the GTA because it is the financial center of the largest city in Canada, likely to become a popular destination for such refugees and homeless, so many new shelters may be built in this Canada's financial center.

Specifically, this paper uses the dataset about occupancy and capacity of daily shelter and overnight service provided by Open Data Toronto, which contains all relevant information of cities in the GTA(Greater Toronto Area). Our estimand, or response variable, is the number of beds needed for a shelter,

The remainder of this paper is structured as follows: @sec-data details the data and measurement process of the chosen dataset; @sec-results covers the multiple linear regression model used in this paper; @sec-results provides results obtain from the derived model; and @sec-discussion includes discussions, such as implications and future steps, of our model, and how our results can in turn be furthermore improved.

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References